---
title: "Untitled"
format: html
editor: visual
---

```{r, load_pckg}

source(here::here('scripts', 'library.R'))
```

```{r load_data}

load(here("data", "data_sample.RData"))
load(here("data", "data_covariate.RData"))
```

# Level: Country-year

## Treatment: High exposure \| severity

```{r}

# define sample
data <- data_sample_iso |> 
  # observation window 2003:2018
  filter(year %in% c(2003:2018)) |> 
  # treatment window 2008:2013 (2003+5:2018-5): include 'clean treated' units + 'never treated' units
  filter(cohort_D_H %in% c(2008:2013, 9999)) |> 
  # treatment: high exposure|severity
  dplyr::select(iso3c, year, t, Y, Y_d, ends_with("_H"), matches("_H_"), gdp:region_agg) |> 
  rename_with(~ str_remove_all(., "D_H_"), contains("D_H_")) |> 
  rename_with(~ str_remove_all(., "_H"), contains("_H"))
  
```

```{r}

## log-transformation
# outcome
data$log_Y <- log(data$Y + .001)

# pre-treatment characteristics
data$log_pop_2003 <- log(data$pop_2003)
data$log_gdp_2003 <- log(data$gdp_2003)
data$log_gdp.pc_2003 <- log(data$gdp.pc_2003)

# label new variables
labelled::var_label(data) <- list(
  log_Y = "ODA commitments, constant 2021 USD (log)",
  log_pop_2003 = "Population in 2003, total (log)",
  pop_0307 = "Change in population 2003-07, total",
  log_gdp_2003 = "GDP in 2003, PPP, constant 2021 international USD (log)",
  gdp_0307 = "Change in GDP 2003-07, PPP, constant 2021 international USD",
  log_gdp.pc_2003 = "GDP per capita in 2003, PPP, constant 2021 international USD (log)",
  gdp.pc_0307 = "Change in GDP per capita 2003-07, PPP, constant 2021 international USD",
  storm_0307	= "Number of storms over the 2003-07 period",
  drought_0307 = "Number of droughts over the 2003-07 period",
  region_agg = "Region",
  inc_gp_03 = "Income group, 2003",
  inc_gp_07 = "Income group, 2007"
)
  
```

```{r}

# Compare units 'ever treated' w/in the *observation* window vs. 'never treated'
# N.B. 'Never-treated' units = never treated w/in the observation window

data |> 
  dplyr::select(iso3c, ever_treated, matches("(gdp|pop|dro|sto|inc|reg).*0(3|7)"), region_agg) |> 
  distinct() |> 
# get_dupes(iso3c)
  mutate(ever_treated = factor(ever_treated, labels = c("Never treated", "Treated (2008-13 cohorts)"))) |> 
  tbl_summary(
    include = c("log_pop_2003", "pop_0307", "log_gdp_2003", "gdp_0307", "log_gdp.pc_2003", "gdp.pc_0307",  "storm_0307", "drought_0307", "region_agg", "inc_gp_03", "inc_gp_07"),
    by = ever_treated,
    statistic = all_continuous() ~ "{mean} ({sd})",
    type = list(c("storm_0307", "drought_0307") ~ "continuous")
  ) |> 
  add_overall(last = TRUE) |> 
  add_p(test = all_continuous() ~ "oneway.test",
        test.args = all_continuous() ~ list(var.equal = TRUE))

```

### Set-up

```{r sample}

# Bin periods outcome time window of interest (-5:5)
data$rel_time <- bin(data$rel_time, list("≤ -6" = ~x < -5, "≥ 6" = ~x > 5))

```

```{r}

visdat::vis_miss(data)

```

```{r}

# Constructs a fixest panel data base
data = panel(data, ~iso3c+year)

# no endogenous treatment timing, i.e.,
# the probability of receiving treatment does not depend on previous outcome dynamics.
```

## Binary and staggered ('first-only') treatment

### TWFE

#### Region-Year FEs

```{r}

# No region-year FEs
twfe <- feols(log_Y ~ i(rel_time, ever_treated, ref = c(-1, 9999)) | iso3c + year, data = data)

# OECD region-year FEs
twfe_reg <- feols(log_Y ~ i(rel_time, ever_treated, ref = c(-1, 9999)) | region_name^year + iso3c + year, data = data)

# Aggregate region-year FEs
twfe_aggreg <- feols(log_Y ~ i(rel_time, ever_treated, ref = c(-1, 9999)) | region_agg^year + iso3c + year, data = data)

# plot
iplot(twfe, x.shift = -.2, col = "grey70", main = "Region-Year FEs")
iplot(twfe_reg, add = TRUE, x.shift = 0, col = "dodgerblue")
iplot(twfe_aggreg, add = TRUE, x.shift = .2, col = "tomato")

legend("topleft", bty = "n",
       c("TWFE", "TWFE | region^year", "TWFE | agg.region^year"),
       pch = c(19, 19, 19),
       col = c("grey70","dodgerblue", "tomato"),
       y.intersp = 0.3)

# NOTE:
# Region-Year FEs 'flatten' estimates 
# Agg. Region CIs < OECD Region CIs
# Updated Baseline = i + t + Agg.Region-Year FEs
```

#### Country-level linear time trends

```{r}

# Baseline + Country-specific LTTs
twfe_aggreg_iso.lt <- feols(log_Y ~ i(rel_time, ever_treated, ref = c(-1, 9999)) | iso3c[t] + region_agg^year + iso3c + year, data = data)

# Baseline + Population_2003 LTTs
twfe_aggreg_pop.lt <- feols(log_Y ~ i(rel_time, ever_treated, ref = c(-1, 9999)) | log_pop_2003[t] +  region_agg^year + iso3c + year, data = data)

# plot
iplot(twfe_aggreg, x.shift = -.2, col = "grey70", main = "Linear time trends")
iplot(twfe_aggreg_iso.lt, add = TRUE, x.shift = 0, col = "dodgerblue")
iplot(twfe_aggreg_pop.lt, add = TRUE, x.shift = .2, col = "tomato")

legend("topleft", bty = "n",
       c("TWFE | agg.region^year (baseline)", "Baseline + Country-LTTs", "Baseline + Pop_2003-LTTs"),
       pch = c(19, 19, 19),
       col = c("grey70","dodgerblue", "tomato"),
       y.intersp = 0.3)

# NOTE:
# Linear trends 'flatten' pre-treat estimates, but similar to post-treat baseline
# Country LTs = Pop_2003 LTs
# LTs CIs < Baseline CIs
# Country LTs CIs < Pop_2003 LTs CIs
# Updated Baseline = i + t + Agg.Region-Year FEs + Country-LTs
```

#### Covariates

##### Population (slow-moving)

```{r}

# Baseline
twfe_aggreg_iso.lt <- feols(log_Y ~ i(rel_time, ever_treated, ref = c(-1, 9999)) | iso3c[t] +  region_agg^year + iso3c + year, data = data)

# Baseline + Population
data$log_pop <- data$
twfe_aggreg_pop <- feols(log_Y ~ l(log_pop, k=-1) + i(rel_time, ever_treated, ref = c(-1, 9999)) | region_agg^year + iso3c + year, data = data)

# Baseline + Population_2003 + GDP_2003
twfe_aggreg_pop_gdp <- feols(log_Y ~ log_pop_2003 + log_gdp_2003 + i(rel_time, ever_treated, ref = c(-1, 9999)) | region_agg^year + iso3c + year, data = data)

# plot
iplot(twfe_aggreg_iso.lt, x.shift = -.2, col = "grey70", main = "Population + GDP (pre-treatment, 2003)")
iplot(twfe_aggreg_pop, add = TRUE, x.shift = 0, col = "dodgerblue")
iplot(twfe_aggreg_pop_gdp, add = TRUE, x.shift = .2, col = "tomato")

legend("topleft", bty = "n",
       c("TWFE + region_agg^year + iso3c[t] (baseline)", "Baseline + Population", "Baseline + Population + GDP"),
       pch = c(19, 19, 19),
       col = c("grey70", "dodgerblue", "tomato"),
       y.intersp = 0.3)

# NOTE:
# Pop. pre-treat estimates are similar to Baseline, but lower post-treat
# Pop. CIs are wider than to Baseline
# No change in Baseline
```

##### Compound events

```{r}

# Baseline + storm
twfe_aggreg_iso.lt_sto <- feols(log_Y ~ storm + i(rel_time, ever_treated, ref = c(-1, 9999)) | iso3c[t] +  region_agg^year + iso3c + year, data = data)

# Baseline + storm + drought
twfe_aggreg_iso.lt_sto_dro <- feols(log_Y ~ storm + drought + i(rel_time, ever_treated, ref = c(-1, 9999)) | iso3c[t] +  region_agg^year + iso3c + year, data = data)


# plot
iplot(twfe_aggreg_iso.lt, x.shift = -.2, col = "grey70", main = "Region-Year FEs")
iplot(twfe_aggreg_iso.lt_sto, add = TRUE, x.shift = 0, col = "dodgerblue")
iplot(twfe_aggreg_iso.lt_sto_dro, add = TRUE, x.shift = .2, col = "tomato")

legend("topleft", bty = "n",
       c("TWFE + region_agg^year + iso3c[t] (baseline)", "Baseline + Storm", "Baseline + storm + drought"),
       pch = c(19, 19, 19),
       col = c("grey70", "dodgerblue", "tomato"),
       y.intersp = 0.3)

# NOTE:
# No change in estimates
# Baseline CIs < Baseline + storm/drought
# No change in Baseline
```

### HTE-Robust TWFE

```{r}

data$iso3 <- as.numeric(factor(data$iso3c))

# estimate group-time average treatment effects without covariates
mw.attgt <- att_gt(yname = "log_Y",
                   gname = "cohort_D",
                   idname = "iso3",
                   tname = "year",
                   xformla = ~1,
                   data = data
                   )

# summarize the results
summary(mw.attgt)

mw.dyn <- aggte(mw.attgt, type = "dynamic")
summary(mw.dyn)

ggdid(mw.dyn)
```

```{r}


est <- lpdid(df = data, window = c(-6, 6),
    y = "log_Y",
    treat_status = "D",
    unit_index = "iso3c", time_index = "year",
    cluster	= "iso3c"
    )

fixef(est)
nlme::fixed.effects(est)

# LP-DID Staggered binary "first treatment"
stag_first_res <- lpdid(df = data, window = c(-10, 10),
    y = "log_Y",
    treat_status = "D_HeHs_first",
    unit_index = "iso3c", time_index = "year",
    cluster	= "iso3c"
    )

# LP-DID Non-absorbing binary treatment (w/out Y_lag)
nabs_res <- lpdid(df = data, window = c(-6, 6),
    y = "log_Y",
    treat_status = "D_HeHs",
    unit_index = "iso3c", time_index = "year",
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    )  

# LP-DID Non-absorbing binary treatment (w/ Y_lag)
nabs_res_l <- lpdid(df = data, window = c(-6, 6),
    y = "log_Y",
    treat_status = "D_HeHs",
    outcome_lags = 1,
    unit_index = "iso3c", time_index = "year",
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    ) 
  
plot_lpdid(stag_first_res, x.shift = -0.1, col = "dodgerblue")
plot_lpdid(nabs_res, x.shift = 0.1, col = "tomato", add = T)
plot_lpdid(nabs_res_l, x.shift = 0.2, col = "orange", add = T)

abline(h = mean(c(15, 15, 35)), lty = 2)
legend("bottomleft", 
       legend = c("Staggered binary first treatment", 
                  "Non-absorbing binary treatment (K=5)",
                  "Non-absorbing binary treatment (K=5) + Y_lag"
                  ), 
       bty = "n",
       col = c("dodgerblue", "tomato", "orange"), pch = 19)
```

## Binary and non-absorbing treatment

### LP-DID

## Heterogeneity

### Treatment intensity

### ODA category

# Level: sector-country-year

## Extensive margin

## Intensive margin

```{r nonabs_lags}

res_1 <- lpdid(df = data, window = c(-7, 10),
    y = "log_Y",
    treat_status = "D_HeHs",
    unit_index = "iso3c", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 1,
    cluster	= "iso3c"
    )
    
res_3 <- lpdid(df = data, window = c(-7, 10),
    y = "log_Y",
    treat_status = "D_HeHs",
    unit_index = "iso3c", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 3,
    cluster	= "iso3c"
    )

res_5 <- lpdid(df = data, window = c(-7, 10),
    y = "log_Y",
    treat_status = "D_HeHs",
    unit_index = "iso3c", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    )

plot_lpdid(res_1, x.shift = -0.1, col = "dodgerblue", main = "Non-absorbing lags")
plot_lpdid(res_3, x.shift = 0, col = "tomato", add = T)
plot_lpdid(res_5, x.shift = 0.1, col = "darkgreen", add = T)

abline(h = mean(c(15, 15, 35)), lty = 2)
legend("topleft", legend = c("K=1", "K=3", "K=5"), bty = "n",
       col = c("dodgerblue", "tomato", "darkgreen"), pch = 19)
```

## Treatment type

```{r}

map(treat_list, \(x) {
  
  res <- lpdid(df = data, window = c(-5, 10),
    y = "log_Y",
    treat_status = x,
    unit_index = "iso3c", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    )

  plot_lpdid(res, x.shift = -0.05, col = "dodgerblue", main = x)

})

```

## Fixed effects

```{r}

FEs <- list(
  ~ 1,
  ~ 1 | region_agg^year,
  ~ 1 | region_code^year,
  ~ 1 | region_code^year + iso3c[[t]]
)
```

```{r}

map(FEs, \(x) {

    res_lpdid <- lpdid(df = data, window = c(-5, 10),
    y = "log_Y",
    treat_status = "D_HeHs",
    unit_index = "iso3c", time_index = "year",
    controls = x,
    outcome_lags = 1,
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    )
    
    titles <- as.character(x)
    
    plot_lpdid(res_lpdid, x.shift = -0.05, col = "dodgerblue",
                       main = titles)
})
```

## Covariates

```{r}

covariates_t <- list(
  ~ sto_dummy,
  ~ storm,
  ~ storm + gdp,
  ~ storm + gdp.pc + pop
)
```

```{r}

map(covariates_t, \(x) {

    res_lpdid <- lpdid(df = data, window = c(-5, 10),
    y = "log_Y",
    treat_status = "D_HeHs",
    unit_index = "iso3c", time_index = "year",
    controls_t = x,
    outcome_lags = 1,
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    )
    
    titles <- sub("^~", "", as.character(x))
    
    plot_lpdid(res_lpdid, x.shift = -0.05, col = "dodgerblue",
                       main = titles)
})
```

# Category-country-year

```{r}

# sample
data <- data_sample_iso_cat |> 
  mutate(
    cat = case_when(
      category == "Development" ~ "dev",
      category == "Gov. & Peace" ~ "gov",
      category == "Budgetary" ~ "budg",
      category == "Humanitarian" ~ "hum"
    ),
    iso_cat = paste0(iso3c, "_", cat)) |> 
  relocate(iso_cat, .after = category)

# Constructs a fixest panel data base
data = panel(data, ~iso_cat+year)

```

```{r}

# outcome
data$log_Y <- log(data$Y + .001)
```

```{r}

# list categories
categories  <- list("Development", "Gov. & Peace", "Budgetary", "Humanitarian")
```

## Unique sample

```{r}

map(treat_list, \(x) {
  
  res <- lpdid(df = data, window = c(-5, 10),
    y = "log_Y",
    treat_status = x,
    unit_index = "iso_cat", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    )

  plot_lpdid(res, x.shift = -0.05, col = "dodgerblue", main = x)
  
})
```

## Split sample

```{r}

map(categories, \(x) {
  
  res <- lpdid(df = data[data$category==x, ], window = c(-5, 10),
    y = "log_Y",
    treat_status = "D_HeHs",
    unit_index = "iso_cat", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    )

  plot_lpdid(res, x.shift = -0.05, col = "dodgerblue", main = x)
  
})
```

# Sector-country-year

```{r}

# sample
data <- data_sample_sect

# Constructs a fixest panel data base
data = panel(data, ~iso_sect_id+year)

# outcome
data$log_Y <- log(data$commit + .001)
```

## Binary outcome (extensive margin)

```{r}

# Binary outcome (Y > 0)
res_1 <- lpdid(df = data, window = c(-5, 10),
    y = "commit_dummy",
    treat_status = "D_HeHs",
    unit_index = "iso_sect_id", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 1,
    cluster	= "iso3c"
    )

res_3 <- lpdid(df = data, window = c(-5, 10),
    y = "commit_dummy",
    treat_status = "D_HeHs",
    unit_index = "iso_sect_id", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 3,
    cluster	= "iso3c"
    )

res_5 <- lpdid(df = data, window = c(-5, 10),
    y = "commit_dummy",
    treat_status = "D_HeHs",
    unit_index = "iso_sect_id", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    )

plot_lpdid(res_1, x.shift = -0.1, col = "dodgerblue", main = "Non-absorbing lags")
plot_lpdid(res_3, x.shift = 0, col = "tomato", add = T)
plot_lpdid(res_5, x.shift = 0.1, col = "darkgreen", add = T)

abline(h = mean(c(15, 15, 35)), lty = 2)
legend("topleft", legend = c("K=1", "K=3", "K=5"), bty = "n",
       col = c("dodgerblue", "tomato", "darkgreen"), pch = 19)
```

## Log-transformed outcome

```{r}

res <- lpdid(df = data, window = c(-5, 10),
    y = "log_Y",
    treat_status = "D_HeHs",
    unit_index = "iso_sect_id", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    )

plot_lpdid(res, x.shift = -0.05, col = "dodgerblue")

```

## Split sample

```{r}

map(categories, \(x) {
  
  res <- lpdid(df = data[data$category==x, ], window = c(-5, 10),
    y = "log_Y",
    treat_status = "D_HeHs",
    unit_index = "iso_sect_id", time_index = "year",
    outcome_lags = 1,
    nonabsorbing_lag = 5,
    cluster	= "iso3c"
    )

  plot_lpdid(res, x.shift = -0.05, col = "dodgerblue", main = x)
  
})
```
