---
title: "Untitled"
format: html
editor: visual
---

```{r, load_pckg}

source(here::here('scripts', 'library.R'))
```

```{r load_data}

load(here("data", "data_sample.RData"))

```

# Model

# Data

Unit: Country \| Time: Year

```{r data}

data <- data_sample_iso |> 
  # observation window 2003:2018
  filter(year %in% c(2003:2018))

## log-transformation

# outcome
data$log_Y <- log(data$Y + .001)

# pre-treatment characteristics
data$log_pop_2003 <- log(data$pop_2003)
data$log_pop_0307 <- log(data$pop_0307)

data$log_gdp_2003 <- log(data$gdp_2003)
data$log_gdp.pc_2003 <- log(data$gdp.pc_2003)

```

```{r}

## label new variables

labelled::var_label(data) <- list(
  log_Y = "ODA commitments, constant 2021 USD (log)",
  log_pop_2003 = "Population in 2003, total (log)",
  pop_0307 = "Change in population 2003-07, total",
  log_gdp_2003 = "GDP in 2003, PPP, constant 2021 international USD (log)",
  gdp_0307 = "Change in GDP 2003-07, PPP, constant 2021 international USD",
  log_gdp.pc_2003 = "GDP per capita in 2003, PPP, constant 2021 international USD (log)",
  gdp.pc_0307 = "Change in GDP per capita 2003-07, PPP, constant 2021 international USD",
  storm_0307	= "Number of storms over the 2003-07 period",
  drought_0307 = "Number of droughts over the 2003-07 period",
  region_agg = "Region",
  inc_gp_03 = "Income group, 2003",
  inc_gp_07 = "Income group, 2007"
)
```

# No Treatment Reversals

## Visualizing Data

```{r}

# Binary high exposure|severity treatment
panelview(1 ~ D_H,                   
            main = "High exposure and/or severity event",          
            by.timing = TRUE,                   
            collapse.history = TRUE,
            pre.post = TRUE,
            data = data,                   
            index = c("iso3c", "year"))


# Binary staggered 'first' high exposure|severity treatment
panelview(1 ~ D_H_first,                   
            main = "Staggered 'first' high exposure|severity treatment",          
            by.timing = TRUE,                   
            collapse.history = TRUE,
            pre.post = TRUE,
            data = data,                   
            index = c("iso3c", "year"))

# binary staggered 'first treatment' treatment status 
treat.vars.first <- c("D_HeLs_first", "D_HeHs_first", "D_LeHs_first")

panel.bin <- treat.vars.first |>      
  map(~ panelview(as.formula(paste("1 ~", .x)),                   
                  main = .x,                   
                  by.timing = TRUE,                   
                  collapse.history = TRUE,
                  pre.post = TRUE,
                  data = data,                   
                  index = c("iso3c", "year")))

```

```{r}

## Relative treatment histories: 
# D_H_first
data |> 
  dplyr::select(iso3c, year, cohort_D_H, D_H_first) |> 
  mutate(rel_time = ifelse(cohort_D_H != 9999, year - cohort_D_H, NA)) |> 
  filter(rel_time %in% c(-5:5)) |> 
  panelview(1 ~ D_H_first,
            by.timing = TRUE,                   
            collapse.history = TRUE,
            pre.post = TRUE,
            main = "Unique Treatment Histories: D_H_first",
            index = c("iso3c", "rel_time"))

# D_He_first
data |> 
  dplyr::select(iso3c, year, cohort_D_He, D_He_first) |> 
  mutate(rel_time = ifelse(cohort_D_He != 9999, year - cohort_D_He, NA)) |> 
  filter(rel_time %in% c(-5:5)) |> 
  panelview(1 ~ D_He_first,
            by.timing = TRUE,                   
            collapse.history = TRUE,
            pre.post = TRUE,
            main = "Unique Treatment Histories: D_He_first",
            index = c("iso3c", "rel_time"))

# D_Hs_first
data |> 
  dplyr::select(iso3c, year, cohort_D_Hs, D_Hs_first) |> 
  mutate(rel_time = ifelse(cohort_D_Hs != 9999, year - cohort_D_Hs, NA)) |> 
  filter(rel_time %in% c(-5:5)) |> 
  panelview(1 ~ D_Hs_first,
            by.timing = TRUE,                   
            collapse.history = TRUE,
            pre.post = TRUE,
            main = "Unique Treatment Histories: D_Hs_first",
            index = c("iso3c", "rel_time"))
```

# Linear models

## TWFE log-OLS

```{r}

# List of treatment variables
treat.vars.first <- c("D_H_first", "D_He_first", "D_Hs_first")

# Iterate over each treatment variable
for (treat_var in treat.vars.first) {
  # Drop always treated units
  df.h <- as.data.frame(data |>
                          group_by(iso3c) |>
                          mutate(treatment_mean = mean(!!sym(treat_var), na.rm = TRUE)))
  df.h.feols <- df.h[which(df.h$treatment_mean < 1), ]

# Baseline specification w/ sub-sample (only 'ever-treated' units)
  formula <- as.formula(paste("log_Y ~", treat_var, "| iso3c[t] + region_agg^year + iso3c + year"))
  res_ever_H <- feols(formula,
                      data = df.h.feols,
                      cluster = "iso3c")

  # Dynamic TWFE
  es.h.feols <- fect::get.cohort(df.h.feols, D = treat_var,
                                  index = c("iso3c", "year"))

  # Drop always treated units
  es.h.feols$treat <- as.numeric(es.h.feols$treatment_mean > 0)
  es.h.feols[which(is.na(es.h.feols$Time_to_Treatment)), 'Time_to_Treatment'] <- 0 # can be an arbitrary value

  # Fit the model
  res_es.h.feols <- feols(log_Y ~ i(Time_to_Treatment, treat, ref = -1) |
                            iso3c[t] + region_agg^year + iso3c + year,
                          data = es.h.feols,
                          cluster = "iso3c")

  # Extract coefficients
  es.h.output <- as.matrix(res_es.h.feols$coeftable)

  # Convert to data frame
  es.h.output <- as.data.frame(es.h.output)

  # Extract the time values from the row names
  time_values <- sub("Time_to_Treatment::([-0-9]+):treat", "\\1", rownames(es.h.output))
  
  # Convert the extracted values to numeric
  time_values <- as.numeric(time_values)
  
  # Add the new variable 'Time' to the data frame
  es.h.output$Time <- time_values+1

  # Add number of treated observations for each time period
  es.h.count <- summarize(es.h.feols,
                          treated_units = sum(treat),
                          .by = Time_to_Treatment)
  es.h.output <- merge(es.h.output, es.h.count, 
                       by.x = "Time", by.y = "Time_to_Treatment")

  # Plot the results
  p.feols.h <- fect::esplot(es.h.output,
                            Period = 'Time', Estimate = 'Estimate',
                            Count = "treated_units", show.count = TRUE,
                            SE = 'Std. Error')
  
  print(p.feols.h)
  
}
```

### Binning

```{r}

# Function to process each treatment variable
bin_es_plot <- function(treat_var) {
  # Drop always treated units
  df.h <- data %>%
    group_by(iso3c) %>%
    mutate(treatment_mean = mean(!!sym(treat_var), na.rm = TRUE)) %>%
    ungroup() %>%
    filter(treatment_mean < 1)

  # Baseline specification w/ sub-sample (only 'ever-treated' units)
  formula <- as.formula(paste("log_Y ~", treat_var, "| iso3c[t] + region_agg^year + iso3c + year"))
  res_ever_H <- feols(formula, data = df.h, cluster = "iso3c")

  # Dynamic TWFE
  es.h.feols <- fect::get.cohort(df.h, D = treat_var, index = c("iso3c", "year"))
  es.h.feols$treat <- as.numeric(es.h.feols$treatment_mean > 0)
  es.h.feols$Time_to_Treatment[is.na(es.h.feols$Time_to_Treatment)] <- 0

  # Bin Time_to_Treatment
  es.h.feols$Time_to_Treatment_bin <- bin(es.h.feols$Time_to_Treatment, list("< -5" = ~x < -5, "> 5" = ~x > 5))

  # Fit the model
  res <- feols(log_Y ~ i(Time_to_Treatment_bin, treat, ref = -1) | iso3c[t] + region_agg^year + iso3c + year,
               data = es.h.feols, cluster = "iso3c")

  # Extract and process coefficients
  output <- as.data.frame(res$coeftable)
  output$Time <- c(-6:-2, 0:6) + 1

  # Add number of treated observations for each time period
  bin_count <- es.h.feols %>%
    group_by(Time_to_Treatment) %>%
    summarize(treated_units = sum(treat), .groups = 'drop')

  output <- merge(output, bin_count, by.x = "Time", by.y = "Time_to_Treatment")

  # Plot the results
  p <- fect::esplot(output, Period = 'Time', Estimate = 'Estimate', SE = 'Std. Error', xlim = c(-6, 6),
                    main = paste("Estimated Binned DTEs:", treat_var))

  # Return the results
  list(output = output, plot = p)
}

# Apply the function to each treatment variable
results <- lapply(treat.vars.first, bin_es_plot)

```

## LP-DID

```{r}


```

## Counterfactual/Imputation estimator

```{r}

df <- data

df$reg_year <- paste0(df$region_agg, "_", df$year)
```

```{r fect}

# Iterate over each treatment variable
for (treat_var in treat.vars.first) {

  # Fit the model
  formula <- as.formula(paste("log_Y ~", treat_var))
  out.fect <- fect(formula, data = df,
                   index = c("iso3c", "year"),
                   method = "cfe", force = "two-way",
                   se = TRUE, parallel = TRUE, nboots = 200,
                   sfe = c("reg_year"),
                   cfe = list(c("iso3c", "t")))

  # Plot the results
  p.fect <- plot(fect.output)
  p.fect

}

```

### Balanced treated sample

Estimate the average treatment effects only for *treated* units that exhibit complete data in specified pre- and post-treatment periods. Here, 5 pre- and post-treatment periods.

```{r fect.balanced}

# Iterate over each treatment variable
for (treat_var in treat.vars.first) {

  # Fit the model
  formula <- as.formula(paste("log_Y ~", treat_var))
  out.fect <- fect(formula, data = df,
                   index = c("iso3c", "year"),
                   method = "cfe", force = "two-way",
                   se = TRUE, parallel = TRUE, nboots = 200,
                   sfe = c("reg_year"),
                   cfe = list(c("iso3c", "t")),
                   balance.period = c(-4,5))

  # Plot the results
  p.fect <- plot(fect.output)
  p.fect

}

```

## Gsynth

```{r}

# Iterate over each treatment variable
for (treat_var in treat.vars.first) {

  # Fit the model
  formula <- as.formula(paste("log_Y ~", treat_var))
  out.fect <- fect(formula, data = df,
                   index = c("iso3c", "year"),
               se = TRUE, method = "gsynth",
               r = 0, CV = FALSE, force = "two-way", 
               nboots = 1000, seed = 02139)

  # Plot the results
  p.fect <- plot(fect.output, type = "gap")
  p.fect
  
}

```

# Non-linear models

## TWFE PPML

```{r twfe.ppml}

# Function to process each treatment variable
bin_es.twfe.ppml_plot <- function(treat_var) {
  # Drop always treated units
  df.h <- data %>%
    group_by(iso3c) %>%
    mutate(treatment_mean = mean(!!sym(treat_var), na.rm = TRUE)) %>%
    ungroup() %>%
    filter(treatment_mean < 1)

  # Dynamic TWFE
  df.h.cohort <- fect::get.cohort(df.h, D = treat_var, index = c("iso3c", "year"))
  df.h.cohort$treat <- as.numeric(df.h.cohort$treatment_mean > 0)
  df.h.cohort$Time_to_Treatment[is.na(df.h.cohort$Time_to_Treatment)] <- 0

  # Bin Time_to_Treatment
  df.h.cohort$Time_to_Treatment_bin <- bin(df.h.cohort$Time_to_Treatment, 
                                           list("< -5" = ~x < -5, "> 5" = ~x > 5))

  # Fit the model
  res <- fepois(Y ~ i(Time_to_Treatment_bin, treat, ref = -1) | 
                       iso3c[t] + region_agg^year + iso3c + year,
               data = df.h.cohort, cluster = "iso3c")

  # Extract and process coefficients
  output <- as.data.frame(res$coeftable)
  output$Time <- c(-6:-2, 0:6) + 1

  # Add number of treated observations for each time period
  bin_count <- df.h.cohort |> 
    group_by(Time_to_Treatment_bin) |> 
    summarize(treated_units = sum(treat), .groups = 'drop')

  output <- merge(output, bin_count, by.x = "Time", by.y = "Time_to_Treatment_bin")

  # Plot the results
  p <- fect::esplot(output, Period = 'Time', 
                    Estimate = 'Estimate', 
                    SE = 'Std. Error', 
                    xlim = c(-6, 6),
                    main = paste("Estimated Binned DTEs:", treat_var))

  # Return the results
  list(output = output, plot = p)
}

# Apply the function to each treatment variable
lapply(treat.vars.first, bin_es.twfe.ppml_plot)

```

## ETWFE

```{r etwfe}

# By default, only use 'not-yet' treated as control group.
# i.e. remove 'always-treated' and 'never-treated'

library(etwfe)

# cohort_D_H
mod_es = etwfe(
  fml  = Y ~ 1, # outcome ~ controls
  tvar = year,        # time variable
  gvar = cohort_D_H, # group variable
  data = data,       # dataset
  vcov = ~iso3c,  # vcov adjustment (here: clustered),
  family = "poisson"
) |> 
  emfx(type = "event")

plot(mod_es)

# cohort_D_Hs
mod_es = etwfe(
  fml  = Y ~ 1, # outcome ~ controls
  tvar = year,        # time variable
  gvar = cohort_D_Hs, # group variable
  data = data,       # dataset
  vcov = ~iso3c,  # vcov adjustment (here: clustered)
  family = "poisson"
) |> 
  emfx(type = "event")

plot(mod_es)

# cohort_D_He
mod_es = etwfe(
  fml  = Y ~ 1, # outcome ~ controls
  tvar = year,        # time variable
  gvar = cohort_D_He, # group variable
# cgroup = "never", # use "never-treated" as control group, return pre-treat efx
  data = data,       # dataset
  vcov = ~iso3c,  # vcov adjustment (here: clustered)
  family = "poisson"
) |> 
  emfx(type = "event")

plot(mod_es)

```

# Heterogeneity

## Disaster magnitude

```{r}

```

## ODA category

```{r Humanitarian}

df.Hum <- data_sample_iso_cat |> 
  filter(category == "Humanitarian") |> 
  dplyr::select(iso3c, year, t, Y, D_H, D_H_first, D_Hs, D_Hs_first, region_agg) |>
  mutate(log_Y = log(Y + .001),
         reg_year = paste0(region_agg, "_", year)
         )

## Non-absorbing
panelview(df.Hum, D = "D_H", index = c("iso3c","year"), 
          by.timing = TRUE)

# Fect
out.fect <- fect(log_Y ~ D_H,
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.Hum,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect, type = "exit", main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

## Absorbing

df.Hum <- df.Hum |> 
  get.cohort(D = "D_H", index = c("iso3c", "year")) |> 
  mutate(
    D_H_short = case_when(
      D_H_first == 1 & (year >= FirstTreat & year < FirstTreat+3) ~ 1,
      TRUE ~ 0),
    D_H_long = case_when(
      D_H_first == 1 & year >= FirstTreat+3 ~ 1,
      TRUE ~ 0)
    )

panelview(df.Hum, D = "D_H_first", index = c("iso3c","year"), 
          by.timing = TRUE)

panelview(df.Hum, D = "D_H_short", index = c("iso3c","year"), 
          by.timing = TRUE)
panelview(df.Hum, D = "D_H_long", index = c("iso3c","year"), 
          by.timing = TRUE)

# Fect
# Short-term [0:3]
out.fect.short <- fect(log_Y ~ D_H_short + D_H_long,
                 D = "D_H_short",
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.Hum,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect.short, main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

# Long-term [>3]
out.fect.long <- fect(log_Y ~ D_H_long + D_H_short,
                 D = "D_H_long",
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.Hum,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect.long, main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

```

```{r Humanitarian.ETWFE}

## Non-linear
# ETWFE

df.Hum.etwfe <- df.Hum |> 
  mutate(first.treat = replace_na(FirstTreat, 0))

library(etwfe)

# cohort_D_H
mod_es = etwfe(
  fml  = Y ~ 1, # outcome ~ controls
  tvar = year, # time variable
  ivar = iso3c, # index variable
  gvar = first.treat, # group variable
  data = df.Hum.etwfe, # dataset
  vcov = ~iso3c,  # vcov adjustment (here: clustered),
  family = "poisson"
) |> 
  emfx(type = "event")

plot(mod_es)

```

```{r Development}

df.Dev <- data_sample_iso_cat |> 
  filter(category == "Development") |> 
  dplyr::select(iso3c, year, t, Y, D_H, D_H_first, D_Hs, D_Hs_first, region_agg) |>
  mutate(log_Y = log(Y + .001),
         reg_year = paste0(region_agg, "_", year)
         )

## Non-absorbing
panelview(df.Dev, D = "D_H", index = c("iso3c","year"), 
          by.timing = TRUE)

# Fect
out.fect <- fect(log_Y ~ D_H,
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.Dev,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect, type = "exit", main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

## Absorbing

df.Dev <- df.Dev |> 
  get.cohort(D = "D_H", index = c("iso3c", "year")) |> 
  mutate(
    D_H_short = case_when(
      D_H_first == 1 & (year >= FirstTreat & year < FirstTreat+3) ~ 1,
      TRUE ~ 0),
    D_H_long = case_when(
      D_H_first == 1 & year >= FirstTreat+3 ~ 1,
      TRUE ~ 0)
    )

# Fect
panelview(df.Dev, D = "D_H_first", index = c("iso3c","year"), 
          by.timing = TRUE)

out.fect.abs <- fect(log_Y ~ D_H_first,
                 D = "D_H_first",
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.Dev,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect.abs, main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

# Short-term [0:3]
panelview(df.Dev, D = "D_H_short", index = c("iso3c","year"), 
          by.timing = TRUE)

out.fect.short <- fect(log_Y ~ D_H_short + D_H_long,
                 D = "D_H_short",
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.Dev,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect.short, main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

# Long-term [>3]
panelview(df.Dev, D = "D_H_long", index = c("iso3c","year"), 
          by.timing = TRUE)

out.fect.long <- fect(log_Y ~ D_H_long + D_H_short,
                 D = "D_H_long",
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.Dev,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect.long, main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")
```

```{r Development.ETWFE}

## Non-linear
# ETWFE

df.Dev.etwfe <- df.Dev |> 
  mutate(first.treat = replace_na(FirstTreat, 0))

library(etwfe)

# cohort_D_H
mod_es = etwfe(
  fml  = Y ~ 1, # outcome ~ controls
  tvar = year, # time variable
  ivar = iso3c, # index variable
  gvar = first.treat, # group variable
  data = df.Dev.etwfe, # dataset
  vcov = ~iso3c,  # vcov adjustment (here: clustered),
  family = "poisson"
) |> 
  emfx(type = "event")

plot(mod_es)

```

## Income group

```{r inc.gp.etwfe}

df.inc <- data |> 
  dplyr::select(iso3c, year, Y, D_H, inc_gp) |> 
  get.cohort(D = "D_H", index = c("iso3c", "year")) |> 
  mutate(first.treat = replace_na(FirstTreat, 0)) |> 
  group_by(iso3c) |>
  mutate(inc.D = replace_na(inc_gp[year == first.treat][1], inc_gp[min(year)]),
         low.inc = case_when(inc.D == "L" ~ TRUE, .default = FALSE)) |>
  ungroup()

mod_es = etwfe(
  fml  = Y ~ 0, # outcome ~ controls
  tvar = year, # time variable
  ivar = iso3c, # index variable
  gvar = first.treat, # group variable
  xvar = low.inc, #  interacted categorical covariate
  data = df.inc, # dataset
  vcov = ~iso3c,  # vcov adjustment (here: clustered),
  family = "poisson"
)

(mod_es_mfx = emfx(mod_es))
emfx(mod_es, hypothesis = "b1 = b2")
```

## State of Emergency

```{r}

# state of emergency
lprop(xtabs(~ D_H + decla_dummy, data))
lprop(xtabs(~ D_Hs + decla_dummy, data))
lprop(xtabs(~ D_He + decla_dummy, data))
lprop(xtabs(~ D_HeHs + decla_dummy, data))
lprop(xtabs(~ D_LeLs + decla_dummy, data))

```

```{r SoE.H}

df.H <- data |> 
  dplyr::select(log_Y, D_H, decla_dummy, region_agg, iso3c, year) |> 
  mutate(
    i.treat = replace_na(ifelse(decla_dummy == 1 & D_H, 1, 0), 0),
    reg_year = paste0(region_agg, "_", year),
    t = year - min(year)
         )

## Non-absorbing
panelview(df.H, D = "i.treat", index = c("iso3c","year"), 
          by.timing = TRUE)

# Fect
out.fect <- fect(log_Y ~ i.treat + D_H + decla_dummy,
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.H,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect, type = "exit", main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

## Absorbing
df.H.first <- get.cohort(df.H, c("iso3c","year"), D = "i.treat") |> 
  mutate(i.treat_first = case_when(
    year >= FirstTreat ~ 1, 
    TRUE ~ 0))

panelview(df.H.first, D = "i.treat_first", index = c("iso3c","year"), 
          by.timing = TRUE, collapse.history = TRUE)

# GSynth
out.gsynth <- fect(log_Y ~ i.treat_first, 
                 index = c("iso3c","year"),
#                 group = "Cohort",
                 data = df.H.first,
                 method = "gsynth", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.gsynth, main = "Estimated ATT (Gsynth)", ylab = "Effect of D on Y", 
  cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

```

```{r SoE.Hs}

df.Hs <- data |> 
  dplyr::select(log_Y, D_Hs, decla_dummy, region_agg, iso3c, year) |> 
  mutate(
    i.treat = replace_na(ifelse(decla_dummy == 1 & D_Hs, 1, 0), 0),
    reg_year = paste0(region_agg, "_", year),
    t = year - min(year)
         )

## Non-absorbing
panelview(df.Hs, D = "i.treat", index = c("iso3c","year"), 
          by.timing = TRUE)

# Fect
out.fect <- fect(log_Y ~ i.treat + D_Hs + decla_dummy, 
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.Hs,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect, type = "exit", main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

## Absorbing
df.Hs.first <- get.cohort(df.Hs, c("iso3c","year"), D = "i.treat") |> 
  mutate(i.treat_first = case_when(
    year >= FirstTreat ~ 1, 
    TRUE ~ 0))

panelview(df.Hs.first, D = "i.treat_first", index = c("iso3c","year"), 
          by.timing = TRUE, collapse.history = TRUE)

# GSynth
out.gsynth <- fect(log_Y ~ i.treat_first, 
                 index = c("iso3c","year"),
#                 group = "Cohort",
                 data = df.Hs.first,
                 method = "gsynth", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.gsynth, main = "Estimated ATT (Gsynth)", ylab = "Effect of D on Y", 
  cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

```

```{r SoE.H.etwfe}

# Absorbing
# NOTE: not enough observations with first treatment x Soe.

# Non-absorbing
data |> 
  dplyr::select(Y, D_H, decla_dummy, iso3c, year) |>
  etwfe(
    fml  = Y ~ 0, # outcome ~ controls
    tvar = year, # time variable
    ivar = iso3c, # index variable
    gvar = D_H, # group variable
    xvar = decla_dummy, #  interacted categorical covariate
    vcov = ~iso3c,  # vcov adjustment (here: clustered),
    family = "poisson"
) |> 
  emfx(hypothesis = "b1 = b2")
```

## Request for international assistance

```{r}

# request for international assistance
lprop(xtabs(~ D_H + appeal_dummy, data))
lprop(xtabs(~ D_Hs + appeal_dummy, data))
lprop(xtabs(~ D_He + appeal_dummy, data))
lprop(xtabs(~ D_HeHs + appeal_dummy, data))
lprop(xtabs(~ D_LeLs + appeal_dummy, data))

```

```{r RIA.H}

df.H <- data |> 
  dplyr::select(log_Y, D_H, appeal_dummy, region_agg, iso3c, year) |> 
  mutate(
    i.treat = replace_na(ifelse(appeal_dummy == 1 & D_H, 1, 0), 0),
    reg_year = paste0(region_agg, "_", year),
    t = year - min(year)
         )

## Non-absorbing
panelview(df.H, D = "i.treat", index = c("iso3c","year"), 
          by.timing = TRUE)

# Fect
out.fect <- fect(log_Y ~ i.treat + D_H + appeal_dummy,
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.H,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect, type = "exit", main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

## Absorbing
df.H.first <- get.cohort(df.H, c("iso3c","year"), D = "i.treat") |> 
  mutate(i.treat_first = case_when(
    year >= FirstTreat ~ 1, 
    TRUE ~ 0))

panelview(df.H.first, D = "i.treat_first", index = c("iso3c","year"), 
          by.timing = TRUE, collapse.history = TRUE)

# GSynth
out.gsynth <- fect(log_Y ~ i.treat_first, 
                 index = c("iso3c","year"),
#                 group = "Cohort",
                 data = df.H.first,
                 method = "gsynth", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.gsynth, main = "Estimated ATT (Gsynth)", ylab = "Effect of D on Y", 
  cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

```

## US Humanitarian response

```{r}

# US Humanitarian response
lprop(xtabs(~ D_H + ofda_bha_dummy, data))
lprop(xtabs(~ D_Hs + ofda_bha_dummy, data))
lprop(xtabs(~ D_He + ofda_bha_dummy, data))
lprop(xtabs(~ D_HeHs + ofda_bha_dummy, data))
lprop(xtabs(~ D_LeLs + ofda_bha_dummy, data))

```

```{r USHum.H}

df.H <- data |> 
  dplyr::select(log_Y, D_H, ofda_bha_dummy, region_agg, iso3c, year) |> 
  mutate(
    i.treat = replace_na(ifelse(ofda_bha_dummy == 1 & D_H, 1, 0), 0),
    reg_year = paste0(region_agg, "_", year),
    t = year - min(year)
         )

## Non-absorbing
panelview(df.H, D = "i.treat", index = c("iso3c","year"), 
          by.timing = TRUE)

# Fect
out.fect <- fect(log_Y ~ i.treat + D_H + ofda_bha_dummy,
                 index = c("iso3c", "year"),
#                 group = "Cohort",
                 data = df.H,
                 method = "fe", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.fect, type = "exit", main = "Estimated ATT", ylab = "Effect of D on Y", cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

## Absorbing
df.H.first <- get.cohort(df.H, c("iso3c","year"), D = "i.treat") |> 
  mutate(i.treat_first = case_when(
    year >= FirstTreat ~ 1, 
    TRUE ~ 0))

panelview(df.H.first, D = "i.treat_first", index = c("iso3c","year"), 
          by.timing = TRUE, collapse.history = TRUE)

# GSynth
out.gsynth <- fect(log_Y ~ i.treat_first, 
                 index = c("iso3c","year"),
#                 group = "Cohort",
                 data = df.H.first,
                 method = "gsynth", force = "two-way",
                 sfe = c("reg_year"), cfe = list(c("iso3c", "t")),
                 se = TRUE, parallel = TRUE, nboots = 200)

plot(out.gsynth, main = "Estimated ATT (Gsynth)", ylab = "Effect of D on Y", 
  cex.main = 0.8, cex.lab = 0.8, cex.axis = 0.8, stats = "F.p")

```
