---
title: "Data: Covariates"
format: html
editor: visual
---

```{r, load_pckg}

source(here::here('scripts', 'library.R'))
```

```{r load_data}

# World Development Indicators (WBG)
wdi_data <- readxl::read_excel("C:\\Users\\pauvernu\\Seafile\\library\\chap_one_data\\source\\covariate\\country_level\\wdi\\P_Data_Extract_From_World_Development_Indicators.xlsx") |> 
  janitor::clean_names() |> 
  pivot_longer(cols = contains("yr"), names_to = "year", values_to = "value") |> 
  mutate(year = as.numeric(str_sub(year, -4, -1)),
         value = as.double(value)) |> 
  filter(!is.na(series_code))
wdi_data$iso3c <- countrycode(sourcevar = wdi_data$country_name,
                             origin = "country.name",
                             destination = "iso3c",
                             warn = T)
wdi_data$iso3c <- ifelse(wdi_data$country_name=="Kosovo", wdi_data$iso3c=="XXK", wdi_data$iso3c)

# UN Population
unpop_data <- read_csv("C:/Users/pauvernu/Seafile/library/chap_one_data/source/covariate/country_level/un_wpp/unpopulation_dataportal_20250623182110.csv") |> 
  janitor::clean_names() |>
  select(country = location, iso3c = iso3, year = time, pop = value)

imf_weo_data <- read.csv("C:\\Users\\pauvernu\\Seafile\\library\\chap_one_data\\source\\covariate\\country_level\\imf\\dataset_2025-06-27T12_48_25.610369931Z_DEFAULT_INTEGRATION_IMF.RES_WEO_6.0.0.csv") |> 
  janitor::clean_names() |> 
  select(country, indicator, starts_with("x")) |> 
  mutate(indicator = ifelse(grepl("GDP", indicator), "gdp_curr_usd", "population")) |> 
  pivot_longer(
    cols = starts_with("x"),
    names_to = "year",
    names_prefix = "x",
    values_to = "value"
  ) |>
  pivot_wider(
    names_from = indicator,
    values_from = value
  ) |> 
  mutate(iso3c = countrycode(sourcevar = country,
                             origin = "country.name",
                             destination = "iso3c",
                             warn = T)) |> 
  mutate(iso3c = case_when(
    grepl("Macao", country) ~ "MAC",
    grepl("Aruba", country) ~ "ABW",
    grepl("Kosovo", country) ~ "XXK",
    grepl("Hong Kong", country) ~ "HKG",
    grepl("European Union (EU)", country) ~ "EUN",
    .default = iso3c
  )) |> 
  relocate(iso3c, .before = country) |> 
  filter(!is.na(iso3c))

# data_crs
load(file = "C:\\Users\\pauvernu\\Seafile\\library\\chap_one_data\\data\\outcome\\data_crs.RData")
```

# **Time-varying**

## Recipient

### GDP

```{r gdp_data}

data_gdp <- wdi_data |> 
  filter(series_name=="GDP, PPP (constant 2021 international $)") |> 
  dplyr::select(iso3c, year, gdp = value)

labelled::var_label(data_gdp) <- list(
  gdp = "GDP, PPP (constant 2021 international $)"
)
```

### Population

```{r pop_data}

data_pop <- unpop_data |> select(-country)

labelled::var_label(data_pop) <- list(
  pop = "Population, total"
)
```

### Other climatic shocks

```{r data_clim_shock}

# create dataset at the country-year level
data_clim_shock <- summarize(data_emdat,
          drought = length(unique(dis_no[grepl("nat-cli-dro-dro", classification_key)])),
          dro_dummy = ifelse(drought > 0, 1, 0),
          storm = length(unique(dis_no[grepl("nat-met-sto", classification_key)])),
          sto_dummy = ifelse(storm > 0, 1, 0),
          .by = c("start_year", "country")
)

# Filter the data for the years between 2003 and 2007 (pre-treatment window)
pre_treat_clim_shock <- subset(data_clim_shock, start_year >= 2003 & start_year <= 2007)
# Estimate the number of droughts, dro_dummy, storms, and sto_dummy by country
sum_pre_treat_clim_shock <- aggregate(
  cbind(drought_0307 = drought, dro_dummy_0307 = dro_dummy, storm_0307 = storm, sto_dummy_0307 = sto_dummy) ~ country, 
  data = pre_treat_clim_shock, FUN = sum)

sum_pre_treat_clim_shock$iso3c <- countrycode(
  sourcevar = sum_pre_treat_clim_shock$country,
  origin = "country.name",
  destination = "iso3c",
  warn = T)
 
data_clim_shock <- sum_pre_treat_clim_shock |> 
  relocate(iso3c, .after = country) |>
  dplyr::select(-country)
```

##### Other non-climatic shocks

##### Request for external assistance/SoE

A request for international assistance from the affected country (`appeal`) and/or a declaration of state of emergency in the country (`declaration`), following a disaster, could affect the probability (and change in value) of nonzero outcome (cf. 'chain' type of causal path).

```{r cov_call}

# create dummy vars whether the affected country released a request for international assistance and/or a SoE declaration  
data_req <- data_emdat |>       
   dplyr::select(dis_no, country, year = start_year, declaration, appeal, ofda_bha = ofda_bha_response) |>       
  mutate(
    declaration = factor(declaration), 
    appeal = factor(appeal),                    
    ofda_bha = factor(ofda_bha),
     ) |>       
   summarize(
     decla_dummy = as.integer(any(declaration == "Yes")),
     appeal_dummy = as.integer(any(appeal == "Yes")),
     ofda_bha_dummy = as.integer(any(ofda_bha == "Yes")),
     .by = c("country", "year")
     )

data_req$iso3c <- countrycode(
  sourcevar = data_req$country,
  origin = "country.name",
  destination = "iso3c",
  warn = T)

labelled::var_label(data_req) <- list(
  decla_dummy = "State of emergency declaration, binary variable",
  appeal_dummy = "Request for international assistance from the affected country, binary variable",
  ofda_bha_dummy = "Response by the (former) Office of US Foreign Disaster Assistance (OFDA) or the Bureau of Humanitarian Assistance (BHA), binary variable"
)

data_req <- data_req |> 
  relocate(iso3c, .after = country) |>
  dplyr::select(-country)
```

### Sector

#### Number of providers per sector-recipient (fragmentation)

```{r cov_sect_frag}

# sum commit by sector-donor-recipient-year 
# for each sector-recipient-year  
# count number of donor with nonzero commit}
```

-   *Share of the largest provider per sector-recipient (concentration)*

```{r cov_sect_conc}

# sum total commit by sector-recipient-year 
# ratio commit / total_commit 
# max ratio by sector-recipient-year}
```

-   *Disbursement rate.*

```{r cov_disb}

# distinct sector-recipient-year 
# sum commit, disb # ratio}
```

-   *Share of commitments through non-State channels.*

# **Time-invariant**

## Recipient

#### *Region*

```{r cov_region}

data_region <-  data_crs |>    
  distinct(iso3c, region_code, region_name) |> 
  mutate(region_agg = case_when(
    region_name %in% c("North of Sahara", "South of Sahara") ~ "Africa",
    region_name %in% c("Caribbean & Central America", "South America") ~ "LAC",
    region_name %in% c("South & Central Asia", "Far East Asia") ~ "Asia",
    .default = region_name
  ))
```

#### *Income groups*

```{r cov_incgp}

data_incgp <- data_crs |>    
  distinct(iso3c, incomegroup_code, incomegroup_name) |> 
  arrange(iso3c)
```

```{r data_covariate}

data_covariate <- data_iso3c_year |> 
#  left_join(data_gdp, by = c("year", "iso3c")) |> 
#  left_join(data_gdp.pc, by = c("year", "iso3c")) |>
  left_join(data_pop, by = c("year", "iso3c")) |>
#  left_join(data_clim_shock, by = "iso3c") |> 
#  left_join(data_req, by = c("year", "iso3c")) |> 
#  left_join(data_incgp_pre, by = "iso3c") |>
  left_join(data_incgp, by = "iso3c") |>
  left_join(data_region, by = "iso3c") |> 
#  mutate(across(drought_0307:sto_dummy_0307, \(x) replace_na(x, 0))) |> 
#  mutate(across(decla_dummy:ofda_bha_dummy, \(x) replace_na(x, 0))) |> 
  arrange(iso3c, year)
```

```{r save_data_covariate}

save(data_covariate, file = "C:\\Users\\pauvernu\\Seafile\\library\\chap_one_data\\data\\covariate\\data_covariate.RData")
```
