---
title: "Data: Outcome"
format: html
editor: visual
---

# Outcome

```{r load}

#load packages
source(here::here('scripts', 'library.R')) 

# load data 
# change path accordingly 
load("C:\\Users\\pauvernu\\Seafile\\library\\chap_one_data\\data\\data_oecd_crs_clean.RData")
```

## Quantity

+ Focus on Official Development Assistance (ODA): Grants (11) and Loans (13)

+ Commitments in constant dollars

> Extensions: consider (i) disbursements; (ii) alternative definitions of nonzero outcomes, based on rounding (ex: at 0.001 MUSD unit) and by including a threshold (ex: \>= 1 000 USD).

```{r data_oda}

data_oda <- data_crs |>
# keep only ODA grants and loans
  filter(flow_code %in% c(11, 13))
```

### Channels

State/Official: Public Sector Institutions
non-State/Official: Multilateral Organizations
non-State/non-Official: NGOs and Civil Society, University, college or other teaching institution, research institute or think‑tank
non-State/non-Official: PPPs and Networks, Private Sector

channel_state = c(10000:12000)
channel_nonstate = c(20000:63000)
channel_nonstate_ngo = c(20000:23000, 51000)
channel_nonstate_multi = c(40000:47000)
channel_nonstate_priv = c(30000:32000, 60000:63000)
channel_other = 90000

### Co-operation modalities

## Unit of observation

Following Faye and Niehaus (2012) and Arezki et al. (2025), I adopt a dyadic approach.

**Advantages**

+ more credible counterfactual. Following a shock, changes in flows received by a recipient can be due to changes in flow from a specific donor or changes in the composition of the pool of donors. remove the risk of composition effect at the donor level (however, it remains at the sector level).

+ allows to distinguish between the effect's extensive vs. intensive margins.

+ allows us to look at the effects of interference (i) between donors, within a recipient country, and (ii) between recipients, within a donor's portfolio.

**Disadvantages**

+ higher risk of measurement error. However, I don't see any systematic bias one way or the other, but this can create noise in the estimation of treatment effects.

+ Faye and Niehaus (2012) and Arezki et al. (2025) don't include multilateral donors, for different reasons. However, I don't see real restriction on including them in the analysis if I use dyad-level FEs.

```{r}
# donors 
length(unique(data_crs$donor_id))
# recipients
length(unique(data_crs$recipient_id))
# pairs
length(unique(data_crs$donor_id)) * length(unique(data_crs$recipient_id))
```
I remove flows to no specific countries, allocated at the regional level.

```{r data_unit}
data_unit <- data_oda |> 
# remove "regional" recipients
  filter(!grepl("regional", recipient_name))
```

> Flows allocated at the regional level represent 8% of total ODA commitments.

```{r sh_reg}

summarize(data_oda,
          regional = sum(usd_commitment_defl[grepl("regional", recipient_name)], na.rm = T),
          total = sum(usd_commitment_defl, na.rm = T),
          reg_share_total = regional/total * 100
          )
```

## Time period

The channel of delivery became an optional reporting item on the new CRS++ reporting scheme in 2004 (cf. Dietrich 2013), so the outcome sample starts at this date. The sample ends in 2022.

```{r data_time}

data_time <- data_unit |> 
  filter(year %in% c(2005:2018))
```

> Note: The treatment sample ends in 2018, the last year available in the GDIS dataset. It results that the final sample is constrained by the outcome sample at its start and by the treatment sample at its end. The final sample covers the period 2005-2018 (13 years).

## Sample

```{r data_sample}

data_sample <- data_time
```

+ **Unit of observation**

```{r unit_outcome}

# define the unit of observation
unit_outcome <- data_sample |> 
  expand(nesting(donor_id, donor_name, cohort_donor), # donor
         nesting(recipient_id, recipient_name, cohort_recipient), # recipient
         year) # year
```

+ **Quantity**

```{r quantity_outcome}

## define the unit-specific quantity
quantity_outcome <- data_sample |> 
  summarize(
# total  
  commit = sum(usd_commitment_defl, na.rm = T),
# channels of delivery 
  commit_nonstate = sum(usd_commitment_defl[channel_nonstate==1], na.rm = T),
  commit_nonstate_ngo = sum(usd_commitment_defl[channel_nonstate_ngo==1], na.rm = T),
  commit_nonstate_multi = sum(usd_commitment_defl[channel_nonstate_multi==1], na.rm = T),
  commit_nonstate_priv = sum(usd_commitment_defl[channel_nonstate_private==1], na.rm = T), 
  .by = c(donor_id, recipient_id, year)
) |> head()
  
 
# co-operation modalities
  commit_budg = sum(usd_commitment_defl[grepl("^A0", aid_t)], na.rm = T),
  commit_multi_core = sum(usd_commitment_defl[aid_t=="B01|B02"], na.rm = T),
  commit_multi_pool = sum(usd_commitment_defl[grepl("^B03", aid_t)], na.rm = T),
  commit_proj = sum(usd_commitment_defl[aid_t=="C01\|D02"], na.rm = T), 
# flow types
  commit_grant = sum(usd_commitment_defl[flow_code==11], na.rm = T),
  commit_loan = sum(usd_commitment_defl[flow_code==13], na.rm = T), 
.by = c(donor_id, recipient_id, year)
)
```

```{r outcome}

## merge outcome data
outcome <- merge(unit_outcome, quantity_outcome, 
                 by=c("donor_id", "recipient_id", "year"), all.x = TRUE) |> 
# treat "true" zeros: NA if country's registration in OECD CRS > current period
  mutate(across(starts_with("commit_"), ~ ifelse(
    is.na(.x) & (cohort_donor <= year | cohort_recipient <= year), 0, .x))) |> 
  # create binary outcome vars
  mutate(commit_dummy = case_when( 
    commit > 0 ~ 1, is.na(commit) ~ NA, .default = 0)
    )
```


## Sectors

+ Remove donor administrative costs, donor country personnel, and debt relief

+ table w/ sectors (row) and channels+modalities (columns)

Add new variable sector_main_code

Table with Total commitments (%) per sectors

Table with Total channels & modalities (%) per sectors

```{r}
# remove Administrative costs not included elsewhere (G01), Donor country personnel (D01), Debt relief (F01)  
  filter(!aid_t %in% c("G01", "D01", "F01")) |> 
  filter(!grepl("^(6\|9)", purpose_code))
```


## Sample

```{r outcome_dupes}

janitor::get_dupes(outcome, purpose_code, recipient_id, year)
```

```{r iso}

# outcome dataset
outcome$iso3c <- countrycode(sourcevar = outcome[, "recipient_name"], 
                             origin = "country.name", destination = "iso3c", warn = T) 
# some values were not matched unambiguously: Kosovo, Micronesia 
outcome <- outcome |> 
  mutate(iso3c = case_when( 
    recipient_name=="Micronesia" ~ "FSM",
    recipient_name=="Kosovo" ~ "XXK", 
    .default = iso3c )
    )
```

```{r}
outcome <- subset(outcome, select = -recipient_id)
```

# Appendix

+ List of donors

+ List of recipients

+ Waffle plot with entry/exit of donors and recipients

+ Table with Total commitments (%) per (i) donor and (ii) recipient

# Save

```{r}

save(outcome, file = here("data", "data_outcome.RData")) rm(delivery_crs_code, data_oda, data_unit, unit_outcome, quantity_outcome)
```